Output Summary

### 2022-01-26 03-post-process/output-plots.Rmd
Add extensive residual diagnostics
And remove old process model residual diagnostics. This is much more detailed now, and is similar to how observation model residuals are diagnosed

## Process Noise Diagnostics

The model assumes mean-zero process noise for all components. That is, the model has an expected value for each component in a given year, and it estimates deviations from this expected value to obtain the realized value in that year. Deviations from the expectation are assumed to be equally likely to be positive or negative for any given year. We can average the process noise components across years to check if it is close to zero. Large deviations from zero in the table below indicate a problem.

```{r get-resids-fn}
# a function to extract the posterior mean of all process model residuals for a single population
get_resids = function(post, j) {

  out = list(
    egg_to_parr_surv = post_summ(post, sub_index("^Lphi_E_Pb_resid[.+,pop]", pop = j))["mean",],
    p_fall_mig = post_summ(post, sub_index("^Lpi_resid[.+,pop]", pop = j))["mean",],
    ow_surv_fall = post_summ(post, sub_index("^Lphi_Pa_Mb_resid[.+,LH_type,pop]", pop = j, LH_type = 1))["mean",],
    ow_surv_spring = post_summ(post, sub_index("^Lphi_Pa_Mb_resid[.+,LH_type,pop]", pop = j, LH_type = 2))["mean",],
    toLGR_surv_nor = post_summ(post, sub_index("^Lphi_Mb_Ma_resid[.+,origin,pop]", pop = j, LH_type = 2, origin = 1))["mean",],
    toLGR_surv_hor = post_summ(post, sub_index("^Lphi_Mb_Ma_resid[.+,origin,pop]", pop = j, LH_type = 2, origin = 2))["mean",],
    LGR_to_BON_surv_nor = post_summ(post, sub_index("^Lphi_Ma_O0_resid[.+,origin]", origin = 1))["mean",],
    LGR_to_BON_surv_hor = post_summ(post, sub_index("^Lphi_Ma_O0_resid[.+,origin]", origin = 2))["mean",],
    yr1_ocean_surv = post_summ(post, sub_index("^Lphi_O0_O1_resid[.+,1,pop]", pop = j))["mean",-1],
    age3_mat_nor = post_summ(post, sub_index("^Lpsi_O1_resid[.+,origin,pop]", origin = 1, pop = j))["mean",],
    age3_mat_hor = post_summ(post, sub_index("^Lpsi_O1_resid[.+,origin,pop]", origin = 2, pop = j))["mean",],
    age4_mat_nor = post_summ(post, sub_index("^Lpsi_O2_resid[.+,origin,pop]", origin = 1, pop = j))["mean",],
    age4_mat_hor = post_summ(post, sub_index("^Lpsi_O2_resid[.+,origin,pop]", origin = 2, pop = j))["mean",],
    BON_to_LGR_surv_nor = post_summ(post, sub_index("^Lphi_Rb_Ra_resid[.+,origin]", origin = 1))["mean",],
    BON_to_LGR_surv_hor = post_summ(post, sub_index("^Lphi_Rb_Ra_resid[.+,origin]", origin = 2))["mean",],
    prespawn_surv = post_summ(post, sub_index("Lphi_Sb_Sa_resid[.+,pop]", pop = j))["mean",]
  )
  out = do.call(cbind, out)
  rownames(out) = NULL
  out
}
```

```{r make-mean-zero-resid-table}
# extract posterior mean residual time series for each population
resids = lapply(1:4, get_resids, post = diag_post)

# convert to array format (list elements become matrix slices)
resids = do.call(abind, append(resids, list(along = 3)))
dimnames(resids)[[3]] = c("CAT", "LOS", "MIN", "UGR")

# calculate the mean of each time series: these should be close to 0
resid_means = t(sapply(1:dim(resids)[2], function(p) colMeans(resids[,p,])))
rownames(resid_means) = dimnames(resids)[[2]]
resid_means = round(resid_means, 3)

# determine the phase of each process model component
is_fw_juvn = rownames(resid_means) %in% c("egg_to_parr_surv", "p_fall_mig", "ow_surv_fall", "ow_surv_spring", "toLGR_surv_nor", "toLGR_surv_hor", "LGR_to_BON_surv_nor", "LGR_to_BON_surv_hor")
is_sw_juvn = stringr::str_detect(rownames(resid_means), "ocean|mat")

# determine the origin of each
is_nor = stringr::str_detect(rownames(resid_means), "nor|egg_to_parr|p_fall|ow_surv")
is_hor = stringr::str_detect(rownames(resid_means), "hor")

# determine the process type: Life History (pi & psi) or Survival (phi)
is_surv = stringr::str_detect(rownames(resid_means), "surv")

# determine the migratory strategy: fall/spring/combined
is_fall = stringr::str_detect(rownames(resid_means), "fall")
is_spring = stringr::str_detect(rownames(resid_means), "spring")

# build ID data frame
resid_IDs = data.frame(
  life_phase = ifelse(is_fw_juvn, "FW Juvenile", ifelse(is_sw_juvn, "SW Juvenile", "FW Adult")),
  type = ifelse(is_surv, "Survival", "Life History"),
  desc = c("Egg-to-Parr", "Strategy Apportionment", "Overwinter", "Overwinter", "Trib to LGR", "Trib to LGR",
           "LGR to BON", "LGR to BON", "Yr1 Ocean", "Age-3 Maturity", "Age-3 Maturity", "Age-4 Maturity", "Age-4 Maturity",
           "BON to LGR", "BON to LGR", "Pre-Spawn"),
  origin = ifelse(is_nor, "NOR", ifelse(is_hor, "HOR", "Combined")), 
  mig_type = ifelse(is_fall, "Fall Migrants", ifelse(is_spring, "Spring Migrants", "Combined"))
)

# combine IDs with mean residual values by population
out = cbind(resid_IDs, resid_means)

# build a kable
kable(out, "html", row.names = FALSE, col.names = c("Life Phase", "Process Type", "Description", "Origin", "Migratory Strategy", colnames(resid_means))) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("condensed")) %>%
  collapse_rows(1:5)
```

@@

>>>>

@@

# Residual Diagnostics {.tabset .tabset-pills}

```{r misc-resid-fns}
get_yr_index = function(params) {
  yr_index = stringr::str_extract(params, "\\[[0-9]+,")
  yr_index = stringr::str_remove(yr_index, "\\[")
  yr_index = stringr::str_remove(yr_index, ",")
  as.numeric(yr_index)
}

make_legend_text = function(mn_post, sd_post, first_last_mn_diff_post) {
  f = function(x) {
    xs = c(mean(x), quantile(x, c(0.025, 0.975)))
    xs = round(xs, 2)
    paste0(xs[1], " (", xs[2], " - ", xs[3], ")")
  }
  c(paste0("Mean: ", f(mn_post)), paste0("SD: ", f(sd_post)), paste0("Diff: ", f(first_last_mn_diff_post)))
}

oma_labels = function(xlab = "", ylab = "", xline = 0.75, yline = 0.75, ...) {
  mtext(side = 1, outer = TRUE, line = xline, text = xlab, ...)
  mtext(side = 2, outer = TRUE, line = yline, text = ylab, ...)
}

no_data_plot = function(main = NULL) {
  plot(1,1, xlim = c(-1,1), ylim = c(-1,1), type = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n", main = main)
  text(0,0, "No Data", font = 3)
}

my_par = function() {
  par(mfrow = c(2,2), mar = c(1,1,2,2), oma = c(2,2,0,0), tcl = -0.15, mgp = c(2,0.35,0))
}
```

#### Why Does this Matter?

This section is devoted to evaluating how consistent the estimated noise terms are with the assumptions of the random processes they are assumed to be generated by.
Specifically, the noise (or "residual") terms should be:

* **Mean-Zero**: each year has an expected value for the realized outcome, and across years errors should cancel out such that the expected residual in any given year is zero.
* **IID**: all residuals are generated by the same random process, i.e., with mean zero and SD equal to one; with the exception of cases with an auto-regressive process included residuals among years are assumed independent.
* **Lack of time series patterns** is implied by these two assumptions, i.e., there should be no time trends or cyclical "regime shifts".

The presence of egregious violations in these assumptions is an indication that the model may give predictions that are inconsistent with the historical data and assumed hypotheses/mechanisms.

#### What was Calculated?

To examine whether these assumptions were consistently met or violated, residuals from the process model and observation model were calculated and standardized.

Standardized residuals were calculated differently depending on the type of random process:

* **Log-normal** (e.g., for abundance data likelihoods): `(log(realized) - log(expected))/SD`, i.e., noise in log space
* **Logit-normal** (e.g., for survival data likelihoods, survival process noise): `(logit(realized) - logit(expected))/SD`, i.e., noise in logit space
* **Binomial and multinomial** (e.g., for carcass count likelihoods): `sqrt(realized) - sqrt(expected)`; for multinomial counts, these were calculated for each category separately, i.e., the Freeman-Tukey statistic, which is similar to the $\chi^2$ statistic but better handles cases where some bin probabilities are small.

For observation model residuals, `SD` represents the assumed observation error variability supplied to the model as known. For process model residuals, `SD` represents the model-estimated `sig` parameter associated with that random process.

This standardization enables comparison across model components that are on different scales (e.g., abundance vs. probability terms) and have different variances, either process or observation.
The specific standardization used implies that the mean of the residuals from a given model component should be zero across all years, should have a standard deviation of 1, and 95% of the residuals should fall between approximately -2 and 2.

#### What is Shown?

For each random process in the model, the time series of standardized noise terms is plotted below (mean and 95% CRI in red, first and last half mean values in blue). To summarize these time series quantitatively, three across-year summary statistics are reported:

* **Across-year mean** -- intended to verify that noise is centered on the expected value overall (ideally equal to 0)
* **Across-year SD** -- intended to verify that noise is as variable as expected overall (ideally equal to 1)
* **mean(last half) - mean(first half)** -- intended to verify that noise is not time-trending (ideally equal to 0)

These values were calculated for each posterior sample and were summarized as the posterior mean and 95% CRI.
They are shown at the top left of each time series plot and summarized across processes in the "Summary Tables" sections.
`r if (params$include_sim_years) "**Although the time series plots show simulated years beyond the observed period, the across-year summary statistics were calculated only using years that have been monitored for at least one population.**"`

## Process Model Noise Terms {.tabset .tabset-pills}

```{r plot-proc-resid-fn}

# function to plot a residual time series for one pop/origin/LH_type/age combo
# returns a data frame with identifiers and across-year summary statistics
plot_proc_resid = function(resid_param, sig_param, year = ts_yrs, pop = NULL, origin = NULL, LH_type = NULL, age = NULL, main = NULL) {

  # extract posterior samples of residual time series
  resid_post = post_subset(post, sub_index(resid_param, year = year, pop = pop, origin = origin, LH_type = LH_type, age = age), matrix = TRUE)

  # extract posterior samples of st. dev. parameter
  sig_post = post_subset(post, sub_index(sig_param, year = year, pop = pop, origin = origin, LH_type = LH_type, age = age), matrix = TRUE)

  # divide each residual by its estimated SD parameterize: to standardize
  resid_post = apply(resid_post, 2, function(x) x/sig_post)

  # extract the year indices available
  y_ind = get_yr_index(colnames(resid_post))
  y_ind_obs = y_ind %in% observable

  # calculate posterior across-year mean and sd for each posterior sample
  # only for years that are observable; don't want to include any simulated years in these stats
  mn_post = apply(resid_post[,y_ind_obs], 1, mean)
  sd_post = apply(resid_post[,y_ind_obs], 1, sd)

  # calculate posterior mean(last half of years) - mean(first half of years)
  y_first = 1:round(length(y_ind[y_ind_obs])/2)
  y_last = (max(y_first) + 1):length(y_ind[y_ind_obs])
  mn_post_first = apply(resid_post[,y_first], 1, mean)
  mn_post_last = apply(resid_post[,y_last], 1, mean)
  first_last_mn_diff_post = mn_post_last - mn_post_first

  # calculate posterior summaries of standardized residual time series
  resid = apply(resid_post, 2, function(x) c(mean = mean(x), sd = sd(x), quantile(x, c(0.5, 0.025, 0.975))))

  # set the y-axis limits: want zero in center and leave room at top for legend
  ylim = max(abs(c(resid["2.5%",], resid["97.5%",])), na.rm = TRUE) * c(-1.4,1.4)

  # blank plot with correct dimensions
  par(lend = "square")
  plot(resid["mean",] ~ all_yrs[y_ind], main = main, type = "n",
       ylim = ylim, xlim = range(all_yrs[y_ind]), xaxt = "n", xlab = "", ylab = "")

  # draw x-axis: years
  at_x = seq(min(all_yrs[y_ind]), max(all_yrs[y_ind]), 4)
  axis(side = 1, at = at_x, labels = paste0("'", substr(at_x, 3, 4)))

  # draw the noise time series: mean and 95% CRI
  polygon(x = c(all_yrs[y_ind], rev(all_yrs[y_ind])),
          y = c(resid["2.5%",], rev(resid["97.5%",])),
          border = NA, col = alpha("salmon", 0.5))
  lines(resid["mean",] ~ all_yrs[y_ind], lwd = 2, col = "red")
  lines(resid["2.5%",] ~ all_yrs[y_ind], lty = 2, col = "red")
  lines(resid["97.5%",] ~ all_yrs[y_ind], lty = 2, col = "red")

  # draw reference line at zero noise
  abline(h = 0, lty = 2, col = "black", lwd = 2)

  # draw posterior mean for first and half of time series
  lines(x = all_yrs[y_ind[y_first]], y = rep(mean(mn_post_first), length(y_first)), col = "blue", lty = 3, lwd = 2)
  lines(x = all_yrs[y_ind[y_last]], y = rep(mean(mn_post_last), length(y_last)), col = "blue", lty = 3, lwd = 2)

  # draw legend that shows estimate summaries
  usr = par("usr"); xdiff = diff(usr[1:2]); ydiff = diff(usr[3:4])
  legend(x = usr[1] - xdiff * 0.05, y = usr[4] + ydiff * 0.01, legend = make_legend_text(mn_post, sd_post, first_last_mn_diff_post), bty = "n", cex = 0.8)

  # create output for building summary tables
  data.frame(param = resid_param,
             pop = ifelse(!is.null(pop), pop, NA),
             origin = ifelse(!is.null(origin), origin, NA),
             LH_type = ifelse(!is.null(LH_type), LH_type, NA),
             age = ifelse(!is.null(age), age, NA),
             n_yrs = ncol(resid_post[,y_ind %in% observable]),
             year_mean_mean = mean(mn_post),
             year_mean_lwr = unname(quantile(mn_post, 0.025)),
             year_mean_upr = unname(quantile(mn_post, 0.975)),
             year_sd_mean = mean(sd_post),
             year_sd_lwr = unname(quantile(sd_post, 0.025)),
             year_sd_upr = unname(quantile(sd_post, 0.975)),
             diff_mean = mean(first_last_mn_diff_post), 
             diff_lwr = unname(quantile(first_last_mn_diff_post, 0.025)), 
             diff_upr = unname(quantile(first_last_mn_diff_post, 0.975))
             )

# container for across-year summary statistics
# is rbinded many times for each random process and pop/origin/LH_type/age combo
proc_error_ests = NULL
```

### FW Juvenile {.tabset .tabset-pills}

#### Egg-to-Parr Survival

```{r egg-to-parr-noise, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lphi_E_Pb_resid[year,pop]", "sig_Lphi_E_Pb[pop]", pop = j, main = pops[j]))
oma_labels("Brood Year", "logit(Egg-to-Parr Survival) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

#### Migratory Strategy Apportionment

```{r pi-noise, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lpi_resid[year,pop]", "sig_Lpi[pop]", pop = j, main = pops[j]))
oma_labels("Brood Year", "logit(Proportion Fall Migrants) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

#### Over-winter Survival {.tabset .tabset-pills}

##### Fall Migrants

```{r overwinter-surv-noise-fall, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lphi_Pa_Mb_resid[year,LH_type,pop]", "sig_Lphi_Pa_Mb[pop,LH_type]", pop = j, LH_type = 1, main = pops[j]))
oma_labels("Brood Year", "logit(Over-winter Survival [Fall Migrants]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

##### Spring Migrants

```{r overwinter-surv-noise-spring, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lphi_Pa_Mb_resid[year,LH_type,pop]", "sig_Lphi_Pa_Mb[pop,LH_type]", pop = j, LH_type = 2, main = pops[j]))
oma_labels("Brood Year", "logit(Over-winter Survival [Spring Migrants]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

#### Trib to LGR {.tabset .tabset-pills}

##### NOR

```{r trib-to-LGR-surv-nor, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lphi_Mb_Ma_resid[year,origin,pop]", "sig_Lphi_Mb_Ma[pop,origin]", pop = j, origin = 1, main = pops[j]))
oma_labels("Brood Year", "logit(Trib to LGR Survival [NOR]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

##### HOR

```{r trib-to-LGR-surv-hor, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lphi_Mb_Ma_resid[year,origin,pop]", "sig_Lphi_Mb_Ma[pop,origin]", pop = j, origin = 2, main = pops[j]))
oma_labels("Brood Year", "logit(Trib to LGR Survival [HOR]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

#### LGR to BON Survival

```{r LGR-to-BON-surv-noise, fig.width = 7, fig.height = 4}
my_par(); par(mfrow = c(1,2))
tmp = lapply(1:jags_data$no, function(o) plot_proc_resid("Lphi_Ma_O0_resid[year,origin]", "sig_Lphi_Ma_O0[origin]", origin = o, main = origins[o]))
oma_labels("Brood Year", "logit(LGR to Ocean Survival) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

### SW Juvenile {.tabset .tabset-pills}

#### Yr1 Ocean Survival

```{r yr1-ocean-surv-noise, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lphi_O0_O1_resid[year,origin,pop]", "sig_Lphi_O0_O1[pop]", origin = 1, year = c(1, ts_yrs), pop = j, main = pops[j]))
oma_labels("Brood Year", "logit(First Year Ocean Survival) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

#### Pr(Mature at age-3) {.tabset .tabset-pills}

##### NOR

```{r pr-mat-3-nor-noise, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lpsi_O1_resid[year,origin,pop]", "sig_Lpsi_O1[pop,origin]", origin = 1, pop = j, main = pops[j]))
oma_labels("Brood Year", "logit(Mature at Age-3 [NOR]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

##### HOR

```{r pr-mat-3-hor-noise, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lpsi_O1_resid[year,origin,pop]", "sig_Lpsi_O1[pop,origin]", origin = 2, pop = j, main = pops[j]))
oma_labels("Brood Year", "logit(Mature at Age-3 [HOR]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

#### Pr(Mature at age-4) {.tabset .tabset-pills}

##### NOR

```{r pr-mat-4-nor-noise, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lpsi_O2_resid[year,origin,pop]", "sig_Lpsi_O2[pop,origin]", origin = 1, pop = j, main = pops[j]))
oma_labels("Brood Year", "logit(Mature at Age-4 [NOR]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

##### HOR

```{r pr-mat-4-hor-noise, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lpsi_O2_resid[year,origin,pop]", "sig_Lpsi_O2[pop,origin]", origin = 2, pop = j, main = pops[j]))
oma_labels("Brood Year", "logit(Mature at Age-4 [HOR]) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

### FW Adult {.tabset .tabset-pills}

#### BON to LGR Survival

```{r BON-to-LGR-surv-noise, fig.width = 7, fig.height = 4}
my_par(); par(mfrow = c(1,2))
tmp = lapply(1:jags_data$no, function(o) plot_proc_resid("Lphi_Rb_Ra_resid[year,origin]", "sig_Lphi_Rb_Ra[origin]", origin = o, main = origins[o]))
oma_labels("Return Year", "logit(BON to LGR Survival) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

#### Pre-spawn Survival

```{r prespawn-surv-noise, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_proc_resid("Lphi_Sb_Sa_resid[year,pop]", "sig_Lphi_Sb_Sa[pop]", pop = j, main = pops[j]))
oma_labels("Return Year", "logit(Pre-spawn Survival) Residual")
proc_error_ests = rbind(proc_error_ests, do.call(rbind, tmp))
```

### **SUMMARY TABLES** {.tabset .tabset-pills}

These tables show posterior summary statistics of the time series of noise terms across years (i.e., the values shown on each figure panel).
Bold values denote that the target value for verifying model assumptions was not within the 95% CRI of the across-year summary statistic and indicates a potential problem.

```{r proc-error-table-prep}
# rename parameters to human-readable format
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_E_Pb_resid\\[year,pop\\]", "Egg to Parr Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lpi_resid\\[year,pop\\]", "Migratory Strategy Apportionment")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Pa_Mb_resid\\[year,LH_type,pop\\]", "Overwinter Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Mb_Ma_resid\\[year,origin,pop\\]", "Trib to LGR Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Ma_O0_resid\\[year,origin\\]", "LGR to BON Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_O0_O1_resid\\[year,origin,pop\\]", "Yr1 Ocean Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lpsi_O1_resid\\[year,origin,pop\\]", "Pr(Mature Age-3)")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lpsi_O2_resid\\[year,origin,pop\\]", "Pr(Mature Age-4)")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Rb_Ra_resid\\[year,origin\\]", "BON to LGR Survival")
proc_error_ests$param = stringr::str_replace(proc_error_ests$param, "Lphi_Sb_Sa_resid\\[year,pop\\]", "Pre-spawn Survival")

# correct origin indices
proc_error_ests$origin[proc_error_ests$param == "Pre-spawn Survival"] = 3
proc_error_ests$origin[is.na(proc_error_ests$origin)] = 1

# correct LH_type indices
proc_error_ests$LH_type[is.na(proc_error_ests$LH_type)] = 3

# correct pop indices
proc_error_ests$pop[is.na(proc_error_ests$pop)] = 5

# remove the age column
proc_error_ests = proc_error_ests[,-which(colnames(proc_error_ests) == "age")]

# replace indices with names
proc_error_ests$pop = c("CAT", "LOS", "MIN", "UGR", "Combined")[proc_error_ests$pop]
proc_error_ests$origin = c("NOR", "HOR", "Combined")[proc_error_ests$origin]
proc_error_ests$LH_type = c("Fall Mig.", "Spring Mig.", "Combined")[proc_error_ests$LH_type]

# which processes are in each life phase?
fw_juv = c("Egg to Parr Survival", "Migratory Strategy Apportionment", "Overwinter Survival", "Trib to LGR Survival", "LGR to BON Survival")
sw_juv = c("Yr1 Ocean Survival", "Pr(Mature Age-3)", "Pr(Mature Age-4)")
fw_adult = c("BON to LGR Survival", "Pre-spawn Survival")

# add a life phase column
proc_error_ests$phase = NA
proc_error_ests$phase = ifelse(proc_error_ests$param %in% fw_juv, 1, proc_error_ests$phase)
proc_error_ests$phase = ifelse(proc_error_ests$param %in% sw_juv, 2, proc_error_ests$phase)
proc_error_ests$phase = ifelse(proc_error_ests$param %in% fw_adult, 3, proc_error_ests$phase)
proc_error_ests$phase = c("FW Juvenile", "SW Juvenile", "FW Adult")[proc_error_ests$phase]

# add levels to each variable
proc_error_ests$phase = factor(proc_error_ests$phase, levels = c("FW Juvenile", "SW Juvenile", "FW Adult"))
proc_error_ests$param = factor(proc_error_ests$param, levels = c(fw_juv, sw_juv, fw_adult))
proc_error_ests$origin = factor(proc_error_ests$origin, levels = c("NOR", "HOR", "Combined"))
proc_error_ests$LH_type = factor(proc_error_ests$LH_type, levels = c("Fall Mig.", "Spring Mig.", "Combined"))
proc_error_ests$pop = factor(proc_error_ests$pop, levels = c("CAT", "LOS", "MIN", "UGR", "Combined"))

# determine whether posterior of across-year quantities overlaps the target value
proc_error_ests$yr_mean_incl_zero = ifelse(proc_error_ests$year_mean_lwr <= 0 & proc_error_ests$year_mean_upr >= 0, TRUE, FALSE)
proc_error_ests$yr_sd_incl_one = ifelse(proc_error_ests$year_sd_lwr <= 1 & proc_error_ests$year_sd_upr >= 1, TRUE, FALSE)
proc_error_ests$diff_incl_zero = ifelse(proc_error_ests$diff_lwr <= 0 & proc_error_ests$diff_upr >= 0, TRUE, FALSE)

# format the mean values to present in table
# round and make bold if target value not in interval
proc_error_ests$year_mean_mean = round(proc_error_ests$year_mean_mean, 2)
proc_error_ests$year_mean_mean = ifelse(!proc_error_ests$yr_mean_incl_zero, paste0("<b>", proc_error_ests$year_mean_mean, "</b>"), proc_error_ests$year_mean_mean)
proc_error_ests$year_sd_mean = round(proc_error_ests$year_sd_mean, 2)
proc_error_ests$year_sd_mean = ifelse(!proc_error_ests$yr_sd_incl_one, paste0("<b>", proc_error_ests$year_sd_mean, "</b>"), proc_error_ests$year_sd_mean)
proc_error_ests$diff_mean = round(proc_error_ests$diff_mean, 2)
proc_error_ests$diff_mean = ifelse(!proc_error_ests$diff_incl_zero, paste0("<b>", proc_error_ests$diff_mean, "</b>"), proc_error_ests$diff_mean)

# drop unneeded columns
proc_error_ests = proc_error_ests[,c("phase", "param", "pop", "origin", "LH_type","year_mean_mean", "year_sd_mean", "diff_mean")]
```

```{r proc-error-table-fn}
proc_error_table = function(var) {
  out = reshape2::dcast(proc_error_ests, phase + param + origin + LH_type ~ pop, value.var = var)
  out[is.na(out)] = ""

  knitr::kable(out, "html", col.names = c("Life Phase", "Process", "Origin", "Migratory Type", "CAT", "LOS", "MIN", "UGR", "Combined"), escape = FALSE, align = "llllccccc") %>%
    kableExtra::kable_styling(full_width = FALSE, bootstrap_options = c("condensed")) %>%
    collapse_rows(columns = 1:4)
}
```

#### Across-Year Mean

The posterior mean of the mean noise value across all years; bold indicates the 95% CRI did not encompass 0.

Values:

* Close to 0 indicate that the time series of noise terms was not biased relative to the expected value (i.e., assumption met)
* Below 0 indicate that the time series of noise terms was biased low relative to the expected value
* Above 0 indicate that the time series of noise terms was biased high relative the expected value

```{r proc-error-mean-table}
proc_error_table("year_mean_mean")
```

#### Across-Year SD

The posterior mean of the SD of noise value across all years; bold indicates the 95% CRI did not encompass 1 (all noise terms were standardized).

Values:

* Close to 1 indicate that the time series of noise terms was as variable as expected (i.e., assumption met)
* Below 1 indicate that the time series of noise terms was less variable than expected
* Above 1 indicate that the time series of noise terms was more variable than expected

```{r proc-error-sd-table}
proc_error_table("year_sd_mean")
```

#### Last Half Mean - First Half Mean

The posterior mean of `mean(noise term in last half of years) - mean(noise term in first half of years)`; bold indicates the 95% CRI did not encompass 0.

Values:

* Close to 0 indicate a lack of an overall time trend (i.e., assumption met)
* Below 0 indicate an overall declining time trend in noise terms
* Above 0 indicate an overall increasing time trend in noise terms

```{r proc-error-diff-table}
proc_error_table("diff_mean")
```

## Observation Model Noise Terms {.tabset .tabset-pills}

```{r plot-obs-resid-fn}

obs_resid_yrs = all_yrs[c(1,observable)]

# function to plot a residual time series for one pop/origin/LH_type/age combo
# returns a data frame with identifiers and across-year summary statistics
plot_obs_resid = function(resid_param, year = ".+", pop = NULL, origin = NULL, LH_type = NULL, age = NULL, main = NULL) {

  # extract posterior samples of residual time series
  resid_post = post_subset(post, sub_index(resid_param, year = year, pop = pop, origin = origin, LH_type = LH_type, age = age), matrix = TRUE)

  # extract the year indices available
  y_ind = get_yr_index(colnames(resid_post))

  # calculate posterior across-year mean and sd for each posterior sample
  mn_post = apply(resid_post, 1, mean)
  sd_post = apply(resid_post, 1, sd)

  # calculate posterior summaries of residual time series
  resid = apply(resid_post, 2, function(x) c(mean = mean(x), sd = sd(x), quantile(x, c(0.5, 0.025, 0.975))))

  # place residuals at correct place in a time series
  # ie., insert missing values
  empty = matrix(NA, nrow = 5, ncol = length(obs_resid_yrs))
  empty[,y_ind] = resid
  resid = empty[,observable]
  rownames(resid) = c("mean", "sd", "50%", "2.5%", "97.5%")
  empty_names = rep(NA, length(obs_resid_yrs))
  empty_names[y_ind] = colnames(resid_post)

  # calculate posterior mean(last half of years) - mean(first half of years)
  y_first = 1:round(length(observable)/2)
  y_last = (max(y_first) + 1):length(observable)
  i_first = as.character(na.omit(empty_names[y_first]))
  i_last = as.character(na.omit(empty_names[y_last]))
  mn_post_first = apply(resid_post[,i_first], 1, mean)
  mn_post_last = apply(resid_post[,i_last], 1, mean)
  first_last_mn_diff_post = mn_post_last - mn_post_first

  # set the y-axis limits: want zero in center and leave room at top for legend
  ylim = max(abs(c(resid["2.5%",], resid["97.5%",])), na.rm = TRUE) * c(-1.4,1.4)

  # blank plot with correct dimensions
  par(lend = "square")
  plot(resid["mean",] ~ all_yrs[observable], main = main,  type = "n", ylim = ylim, xlim = range(all_yrs[observable]), xaxt = "n", xlab = "", ylab = "")

  # draw x-axis: years
  at_x = seq(min(all_yrs[observable]), max(all_yrs[observable]), 4)
  axis(side = 1, at = at_x, labels = paste0("'", substr(at_x, 3, 4)))

  # draw the noise time series: mean and 95% CRI
  segments(all_yrs[observable], resid["2.5%",], all_yrs[observable], resid["97.5%",], col = "red")
  points(resid["mean",] ~ all_yrs[observable], pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)

  # draw reference line at zero noise
  abline(h = 0, lty = 2, col = "black", lwd = 2)

  # draw posterior mean for first and half of time series
  lines(x = all_yrs[observable[y_first]], y = rep(mean(mn_post_first), length(y_first)), col = "blue", lty = 3, lwd = 2)
  lines(x = all_yrs[observable[y_last]], y = rep(mean(mn_post_last), length(y_last)), col = "blue", lty = 3, lwd = 2)

  # draw legend that shows estimate summaries
  usr = par("usr"); xdiff = diff(usr[1:2]); ydiff = diff(usr[3:4])
  legend(x = usr[1] - xdiff * 0.05, y = usr[4] + ydiff * 0.01, legend = make_legend_text(mn_post, sd_post, first_last_mn_diff_post), bty = "n", cex = 0.8)

  # create output for building summary tables
  data.frame(param = resid_param,
             pop = ifelse(!is.null(pop), pop, NA),
             origin = ifelse(!is.null(origin), origin, NA),
             LH_type = ifelse(!is.null(LH_type), LH_type, NA),
             age = ifelse(!is.null(age), age, NA),
             n_yrs = ncol(resid_post[,y_ind %in% observable]),
             year_mean_mean = mean(mn_post),
             year_mean_lwr = unname(quantile(mn_post, 0.025)),
             year_mean_upr = unname(quantile(mn_post, 0.975)),
             year_sd_mean = mean(sd_post),
             year_sd_lwr = unname(quantile(sd_post, 0.025)),
             year_sd_upr = unname(quantile(sd_post, 0.975)),
             diff_mean = mean(first_last_mn_diff_post), 
             diff_lwr = unname(quantile(first_last_mn_diff_post, 0.025)), 
             diff_upr = unname(quantile(first_last_mn_diff_post, 0.975))
  )
}

# container for across-year summary statistics
# is rbinded many times for each random process and pop/origin/LH_type/age combo
obs_error_ests = NULL
```

### Abundance Data {.tabset .tabset-pills}

#### Fall Parr Abundance

```{r fall-trap-resid, fig.width = 6, fig.height = 5}
my_par()
tmp = lapply(1:jags_data$nj, function(j) plot_obs_resid("lPa_obs_resid[year,LH_type,pop]", LH_type = 1, pop = j, main = pops[j]))
oma_labels("Brood Year", "log(Fall Parr Abundance) Residual")
obs_error_ests = rbind(obs_error_ests, do.call(rbind, tmp))
```

#### Spring Smolt Abundance
